---
title: "Analyzing heterogeneous effects of loss-versus-gain framing with causal machine learning"
output:
  html_document:
    df_print: paged
---

# Initialize notebook

Load required package and versions used in this paper

```{r load packages, warning=FALSE, message=FALSE}
library(tidyverse)       # tidyverse_2.0.0
library(vtable)          # vtable_1.4.8 
library(cowplot)         # cowplot_1.1.3
library(stargazer)       # stargazer_5.2.3 
library(grf)             # grf_2.4.0 
library(policytree)      # policytree_1.2.3
library(maq)             # maq_0.6.0
library(sampleSelection) # sampleSelection_1.2-12
library(cobalt)          # cobalt_4.6.1
library(emmeans)         # emmeans_1.11.2
library(tools)           # base package
library(stringr)         # stringr_1.5.1 
library(fastshap)        # fastshap_0.1.1 
library(shapviz)         # shapviz_0.10.1  
library(mgcv)            # mgcv_1.9-1 
```

Set up workspace

```{r setup}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
rm(list=ls())
set.seed(42)
options(scipen=10000)
```

Constants that are used through out the script

```{r}
relevant_features = c(
  "gender",
  "insurance_holder_age",
  "has_company_pension_plan",
  "contract_valuation",
  "historical_payments_cnt", 
  "handling_and_moving", 
  "information_skills", 
  "working_with_computers", 
  "constructing", 
  "management_skills",
  "working_with_machinery_and_specialised_equipment",
  "assisting_and_caring",
  "communication__collaboration_and_creativity"
  )
```

Helper function based on: https://grf-labs.github.io/grf/articles/ate_transport.html

```{r}
average_treatment_effect_test <- function(forest, X.test, p.hat = NULL) {
  Y.hat <- forest$Y.hat
  Y.orig <- forest$Y.orig
  W.hat <- forest$W.hat
  W.orig <- forest$W.orig
  n.rct <- length(Y.orig)
  n.obs <- nrow(X.test)

  if (is.null(p.hat)) {
    S.forest <- regression_forest(X = rbind(forest$X.orig, X.test),
                                  Y = c(rep(1, n.rct), rep(0, n.obs)),
                                  num.trees = 500)
    p.hat <- predict(S.forest)$predictions[1:n.rct]
  }

  tau.hat.test <- predict(forest, X.test)$predictions
  tau.hat.train <- predict(forest)$predictions
  debiasing.weights <- (1 - p.hat) / p.hat * (W.orig - W.hat) / (W.hat * (1 - W.hat))
  Y.residual <- Y.orig - (Y.hat + tau.hat.train * (W.orig - W.hat))

  tau.hat <- mean(tau.hat.test) + sum(debiasing.weights * Y.residual) / n.obs
  sigma2.hat <- var(tau.hat.test) / n.obs + sum(Y.residual^2 * (debiasing.weights / n.obs)^2)

  c(estimate = tau.hat, std.err = sqrt(sigma2.hat))
}
```

# Load Data

```{r}
data_started <- read_csv('data_started.csv')
data_finished <- read_csv('data_finished.csv')
```


# Summary Stats

Number of participants in the two studies for those who made a payment and for all

```{r}
data_started %>% 
  group_by(year) %>% 
  summarise(n = n())

data_finished %>% 
  group_by(year) %>% 
  summarise(n = n())
```

Overall conversion rate (i.e., those who made a payment vs. all) along conditions and experiments

```{r}
data_started %>% 
  summarise(
    avg_conversion = mean(converted)
  )
```

```{r}
data_started %>% 
  group_by(year) %>% 
  summarise(
    avg_conversion = mean(converted)
  )

data_started %>% 
  group_by(year, condition_3lvl) %>% 
  summarise(
    avg_conversion = mean(converted)
  )
```

Overall payment for those who made a payment along conditions and experiments

```{r}
data_finished %>%
  summarise(n = n(), 
            mean_amount = mean(paid_amount),
            median_amount = median(paid_amount),
            sd_amount = sd(paid_amount))

```

```{r}
data_finished %>% 
  group_by(year) %>% 
  summarise(n = n(), 
            mean_amount = mean(paid_amount),
            median_amount = median(paid_amount),
            sd_amount = sd(paid_amount))

data_finished %>% 
  group_by(year, condition_3lvl) %>% 
  summarise(n = n(), 
            mean_amount = mean(paid_amount),
            median_amount = median(paid_amount),
            sd_amount = sd(paid_amount))
```

Covariate summary stats and randomization checks

```{r}
st(data_started, group = 'condition', vars = relevant_features, group.test = TRUE)
```

# Causal ML

## DV: Conversion

```{r}
rm(list=setdiff(ls(), c("relevant_features", "split_ratio", "average_treatment_effect_test", "data_started", "data_finished")))
set.seed(42)
options(scipen=10000)
```

### Fit Causal Forest

Split data into training and test set

```{r}
train <- data_started %>% filter(year==2021)
test <- data_started%>% filter(year==2022)
```

Split data into X, Y, W

```{r}
Y.train <- train$converted
W.train <- ifelse(train$condition_3lvl=='loss', 1, 0)
X.train <- train[relevant_features]

Y.test <- test$converted
W.test <- ifelse(test$condition_3lvl=='loss', 1, 0)
X.test <- test[relevant_features]
```

Fit or load Model

```{r}
if (file.exists("cf_train_conversion.rds")) {
  print("Load CF Train")
  cf.train <- readRDS(file = "cf_train_conversion.rds")
} else {
  print("Fit CF Train")
  cf.train <- causal_forest(X = X.train, Y = Y.train, W = W.train,
                            honesty = FALSE,
                            seed = 42)
  saveRDS(cf.train, file = "cf_train_conversion.rds")
}
```

###  ATE

```{r}
ate <- average_treatment_effect_test(cf.train, X.test)
print(ate[1])
print(ate[1]-1.96*ate[2])
print(ate[1]+1.96*ate[2])
```

###  CATE

```{r}
cate.test <- predict(cf.train, X.test, estimate.variance = TRUE)
cate.test <- cate.test %>%
  mutate(
    cate_direction = ifelse(cate.test$predictions>=0, 'positive', 'negative') 
  )

summary(cate.test$predictions)
```

Distribution of CATE

```{r}
cate_p1 <- ggplot(cate.test, aes(x = predictions, fill = cate_direction)) +
  geom_histogram(binwidth = 0.01, color = "white", boundary  = 0) +
  scale_fill_manual(name = "Direction", values = c(positive = "#1F77B4", negative = "#A56CBD")) +
  scale_x_continuous(labels = scales::label_percent(accuracy = 0.01, suffix = " %")) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(x = "CATE (in percent)", y = "Frequency") +
  geom_vline(xintercept=ate[1]) +
  geom_vline(xintercept=c(ate[1]-1.96*ate[2],ate[1]+1.96*ate[2]), linetype="dashed") +
  theme_bw() + 
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12)) +
  theme(legend.position = "none") +
  annotate("text", x = 0.15, y = 3000, 
           label = sprintf("ATE: %.3f%% [95%% CI: %.3f%%, %.3f%%]", 
                           ate[1]*100, 
                           (ate[1]-1.96*ate[2])*100, 
                           (ate[1]+1.96*ate[2])*100)) +
  
  annotate("segment", x = 0.10, xend = 0.20, y = 2000, yend = 2000, 
           arrow = arrow(length = unit(0.3, "cm")), linewidth = 1) +
  
  annotate("segment", x = -0.10, xend = -0.20, y = 2000, yend = 2000, 
           arrow = arrow(length = unit(0.3, "cm")), linewidth = 1) +
  
  annotate("text", x = 0.15, y = 1300, 
           label = "Respond positively\nto loss-framing", size = 4) +

  annotate("text", x = -0.15, y = 1300, 
           label = "Respond positively\nto gain-framing", size = 4) +
  
  coord_cartesian(xlim = c(-0.25, 0.25))

cate_p1
```

### RATE

```{r}
if (file.exists("cf_test_conversion.rds")) {
  print("Load CF Test")
  cf.test <- readRDS(file = "cf_test_conversion.rds")
} else {
  print("Fit CF Test")
  cf.test <- causal_forest(X = X.test, Y = Y.test, W = W.test,
                          honesty = FALSE,
                          seed = 42)
  saveRDS(cf.test, file = "cf_test_conversion.rds")
}
```

```{r}
rate <- rank_average_treatment_effect(cf.test, cate.test$predictions)
rate$estimate
rate$estimate-1.96*rate$std.err
rate$estimate+1.96*rate$std.err
```

Targeting Operator Characteristic (TOC) for the CATE-based policy with conversion as the dependent variable

```{r}
ggplot() +
  geom_line(mapping = aes(x = q, y = estimate), 
            data=rate$TOC, 
            color = "#1F77B4", size = 1) +
  geom_ribbon(mapping = aes(x = q, y = estimate, ymin = estimate-1.96*std.err, ymax = estimate+1.96*std.err),
              data=rate$TOC, 
              fill = "#1F77B4", alpha = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 1, by = 0.1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.01, suffix = " %")) +
  labs(title = "Targeting Operator Characteristic", x = "Treated Fraction", y = "") +
  theme_bw() + 
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12))

ggsave("RATE - Conversion.pdf", width = 10, height = 8)
```


### GATE

GATE by CATE sign (+/-)

```{r}
X.test_tmp <- X.test
X.test_tmp$cate <- cate.test$predictions

X.test_cate_pos <- X.test_tmp %>% 
  filter(cate > 0) %>% 
  select(-cate)

X.test_cate_neg <- X.test_tmp %>% 
  filter(cate <= 0) %>% 
  select(-cate)

cate_pos <- average_treatment_effect_test(cf.train, X.test_cate_pos)
cate_neg <- average_treatment_effect_test(cf.train, X.test_cate_neg)

cate_sign_df <- data.frame(
  cate_sign = c("Positive", "Negative"),
  estimate = c(cate_pos[1], cate_neg[1]),
  se = c(cate_pos[2], cate_neg[2]),
  n = c(nrow(X.test_cate_pos), nrow(X.test_cate_neg))
)

cate_sign_df <- cate_sign_df %>% arrange(desc(cate_sign))

gate_p1 <- ggplot(data = cate_sign_df) +
  geom_col(aes(x = cate_sign, y = estimate, fill = cate_sign)) +
  scale_fill_manual(values = c(Positive = "#1F77B4", Negative = "#A56CBD")) + 
  geom_errorbar(aes(x = cate_sign, ymin = estimate - 1.96 * se, ymax = estimate + 1.96 * se), width = 0.1) +
  geom_text(aes(x = 1, y = -0.01), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(cate_sign_df[2,"n"]))) +
  geom_text(aes(x = 2, y = 0.01), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(cate_sign_df[1,"n"]))) +
  geom_text(aes(x = 1, y = -0.065), label = paste0(round(cate_sign_df[2,"estimate"]*100,3), " %")) +
  geom_text(aes(x = 2, y = 0.065), label = paste0(round(cate_sign_df[1,"estimate"]*100,3), " %")) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.01, suffix = " %")) +
  ylab("Group ATE") +
  xlab("Causal ML Policy") +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("Negative" = "Not targeted", "Positive" = "Targeted")) +
  theme_bw() +
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12)) +
  theme(legend.position = "none") 

gate_p1
```

GATE by CATE quartiles

```{r}
X.test_tmp <- X.test
X.test_tmp$cate <- cate.test$predictions
X.test_tmp$cate_quartile <- as.factor(ntile(X.test_tmp$cate, 4))

X.test_Q1 <- X.test_tmp %>% 
  filter(cate_quartile == "1") %>% 
  select(-cate,-cate_quartile)

X.test_Q2 <- X.test_tmp %>% 
  filter(cate_quartile == "2") %>% 
  select(-cate,-cate_quartile)

X.test_Q3 <- X.test_tmp %>% 
  filter(cate_quartile == "3") %>% 
  select(-cate,-cate_quartile)

X.test_Q4 <- X.test_tmp %>% 
  filter(cate_quartile == "4") %>% 
  select(-cate,-cate_quartile)

gate_q1 <- average_treatment_effect_test(cf.train, X.test_Q1)
gate_q2 <- average_treatment_effect_test(cf.train, X.test_Q2)
gate_q3 <- average_treatment_effect_test(cf.train, X.test_Q3)
gate_q4 <- average_treatment_effect_test(cf.train, X.test_Q4)

gate_quartiles_df <- data.frame(
  quartile = c("Q1", "Q2", "Q3", "Q4"),
  estimate = c(gate_q1[1], gate_q2[1], gate_q3[1], gate_q4[1]),
  se = c(gate_q1[2], gate_q2[2], gate_q3[2], gate_q4[2]),
  sign = c(as.factor(sign(gate_q1[1])), as.factor(sign(gate_q2[1])), as.factor(sign(gate_q3[1])), as.factor(sign(gate_q4[1]))),
  n = c(nrow(X.test_Q1), nrow(X.test_Q2), nrow(X.test_Q3), nrow(X.test_Q4))
)

gate_p2 <- ggplot(data = gate_quartiles_df) +
  geom_col(aes(x = quartile, y = estimate, fill = sign)) +
  scale_fill_manual(values = c("-1" = "#A56CBD", "1" = "#1F77B4")) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.01, suffix = " %")) +
  geom_errorbar(aes(x = quartile, ymin = estimate - 1.96 * se, ymax = estimate + 1.96 * se), width = 0.1) +
  #ylim(-0.110, 0.110) + 
  ylab("Group ATE") +
  xlab("CATE Quartile") +
  theme_bw() +
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12)) +
  theme(legend.position = "none") 

gate_p2
```

```{r}
plot_grid(cate_p1, gate_p2, gate_p1, labels = c('a', 'b', 'c'), 
          ncol = 1, align = 'h', 
          scale = c(.9, .9, .9)
          )

ggsave("CATE_GATE - Conversion.pdf", width = 10, height = 8)
```

### XAI

Variable importance in causal forest model (conversion)

```{r}
vi <- variable_importance(cf.train)
vi_df <- data.frame(vi)
vi_df$variable <- colnames(cf.train$X.orig)

vi_df <- vi_df %>% 
  select("variable", "vi") %>% 
  arrange(desc(vi))

xai_p1 <- ggplot(vi_df, aes(x = reorder(variable, vi), y = vi)) +
  geom_bar(stat = "identity", fill = "#1F77B4") +
  coord_flip() +
  labs(x = "Covariate", y = "Variable Importance") +
  scale_x_discrete(
    labels = function(x) toTitleCase(gsub("_", " ", x))
  ) +
  theme_bw() + 
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12))

xai_p1
```

Partial dependence plots for causal forest mode (conversion)

```{r}
p_list <- list()
for (var in vi_df$variable) {

  if (var == "gender" | var == "has_company_pension_plan") {
    range <- seq(0, 1, by = 1)
  } else {
    range <- seq(min(X.test[[var]]), max(X.test[[var]]), length.out = 4)    
  }

  range <- seq(min(X.test[[var]]), max(X.test[[var]]), length.out = 4)
  
  results <- lapply(range, function(val) {
    X_mod <- X.test
    X_mod[[var]] <- val
    
    preds <- predict(cf.train, X_mod, estimate.variance = TRUE)
    
    pred_point <- mean(preds$predictions)
    pred_var_mean <- mean(preds$variance.estimates) / nrow(X_mod)
    pred_se <- sqrt(pred_var_mean)
    
    pred_lb <- pred_point - 1.96 * pred_se
    pred_ub <- pred_point + 1.96 * pred_se
    
    data.frame(
      working_with_computers = val,
      pred_point = pred_point,
      pred_lb = pred_lb,
      pred_ub = pred_ub
    )
  })
  
  pdp_df <- do.call(rbind, results)
  names(pdp_df)[1] <- "value"

  p <- ggplot(pdp_df, aes(x = value, y = pred_point)) +
    geom_line(color = "#1F77B4", size = 1) +
    geom_ribbon(aes(ymin = pred_lb, ymax = pred_ub), fill = "#1F77B4", alpha = 0.3) +
    scale_y_continuous(labels = scales::label_percent(accuracy = 0.01, suffix = " %")) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
    labs(x = str_wrap(toTitleCase(gsub("_", " ", var)), width = 12), y = "CATE") +
    #ylim(y = c(-500, 1200)) +
    theme_bw() + 
    theme(text = element_text(size = 10), axis.text.x=element_text(size=8), axis.text.y=element_text(size=8))

  p_list[[var]] <- p
  
}

p_grid <- plot_grid(plotlist = p_list, ncol = 5, align = 'hv')
p_grid
```

```{r}
plot_grid(
  xai_p1, 
  p_grid,
  labels = c('a', 'b'),
  ncol = 1,
  align = "h",
  scale = c(.9, .9),
  rel_heights = c(0.35, 0.65) 
)

ggsave("XAI PDP - Conversion Combined.pdf", width = 10, height = 12)

```

SHAP

```{r}
# Prediction wrapper for grf causal forest
pred_fun <- function(object, newdata, ...) {
  predict(object, newdata, ...)$predictions
}

# Compute SHAP values with fastshap
fs <- fastshap::explain(
  object       = cf.train,
  X            = X.test,
  pred_wrapper = pred_fun,
  nsim = 10
)

# Turn fastshap result into a shapviz object
shap_values <- shapviz(fs, X = X.test)

sv_importance(shap_values, kind = "bee")
ggsave("SHAP - Beeswarm Conversion.png", width = 10, height = 8)

```

### Economic Impact

Data preparation

```{r}
X.test_tmp <- X.test
X.test_tmp$cate <- cate.test$predictions
X.test_tmp$condition <- W.test
X.test_tmp$converted <- Y.test
```

#### Random assignment

Conversion rate under randomized assignment

```{r}
X.test_tmp %>% 
  summarise(n = n(), mean_converted = mean(converted))
```

#### All loss

Conversion rate when all participants are assigned to loss-framing

```{r}
X.test_tmp %>% 
  filter(condition == 1) %>%
  summarise(n = n(), mean_converted = mean(converted))
```

#### All gain

Conversion rate when all participants are assigned to gain-framing

```{r}
X.test_tmp %>% 
  filter(condition == 0) %>%
  summarise(n = n(), mean_converted = mean(converted))
```

#### By CATE sign (+/-)

Conversion rate for optimal assignment to gain- vs. loss-framing

```{r}
# Received gain message and have positive CATE (i.e., should have received a loss message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(cate > 0) %>%
  summarise(n = n(), 
            base = mean(converted),
            policy_cate = base + cate_pos[1]) # estimated conversion rate if they had received a loss message instead

# Received gain message and have negative CATE (i.e., should have received a gain message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(cate <= 0) %>%
  summarise(n = n(), 
            base = mean(converted),
            policy_cate = base # nothing to add because they already got the right message
            )
```

#### By gender

```{r}

X.test_tmp <- X.test

X.test_gender_male <- X.test_tmp %>%
  filter(gender == 0)

X.test_gender_female <- X.test_tmp %>%
  filter(gender == 1)

gender_male <- average_treatment_effect_test(cf.train, X.test_gender_male)
gender_female <- average_treatment_effect_test(cf.train, X.test_gender_female)

gender_df <- data.frame(
  gender = c("Male", "Female"),
  estimate = c(gender_male[1], gender_female[1]),
  se = c(gender_male[2], gender_female[2]),
  n = c(nrow(X.test_gender_male), nrow(X.test_gender_female))
)

gender_df <- gender_df %>% arrange(desc(gender))

gate_p3 <- ggplot(data = gender_df) +
  geom_col(aes(x = gender, y = estimate, fill = gender)) +
  scale_fill_manual(values = c(Male = "#1F77B4", Female = "#A56CBD")) +
  geom_errorbar(aes(x = gender, ymin = estimate - 1.96 * se, ymax = estimate + 1.96 * se), width = 0.1) +
  geom_text(aes(x = 1, y = -0.015), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(gender_df[2,"n"]))) +
  geom_text(aes(x = 2, y = -0.015), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(gender_df[1,"n"]))) +
  geom_text(aes(x = 1, y = 0.02), label = paste0(round(gender_df[2,"estimate"]*100,3), " %")) + # NEW
  geom_text(aes(x = 2, y = 0.02), label = paste0(round(gender_df[1,"estimate"]*100,3), " %")) + # NEW
  ylab("Group ATE") +
  xlab("Gender") +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("Negative" = expression(pi(x[i]) == 0), "Positive" = expression(pi(x[i]) == 1))) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.01, suffix = " %")) + # NEW
  theme_bw() +
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12)) +
  theme(legend.position = "none")

gate_p3

ggsave("GATE by Gender - Conversion.pdf", width = 10, height = 8)
```

```{r}
# female -> loss
# male -> gain

X.test_tmp$cate <- cate.test$predictions
X.test_tmp$condition <- W.test
X.test_tmp$converted <- Y.test

# Received gain message and are male (i.e., should have received a gain message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(gender == 0) %>%
  summarise(n = n(), 
            gender = mean(gender),
            base = mean(converted),
            policy_cate = base) # nothing to add because they already got the right message

# Received gain message and are female (i.e., should have received a loss message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(gender == 1) %>%
  summarise(n = n(), 
            gender = mean(gender),
            base = mean(converted),
            policy_cate = base + gender_female[1] # estimated conversion rate if they had received a loss message instead
            )
```

```{r}
rate <- rank_average_treatment_effect(cf.test, X.test_tmp$gender)
rate$estimate
rate$estimate-1.96*rate$std.err
rate$estimate+1.96*rate$std.err

ggplot() +
  geom_line(mapping = aes(x = q, y = estimate), 
            data=rate$TOC, 
            color = "#1F77B4", size = 1) +
  geom_ribbon(mapping = aes(x = q, y = estimate, ymin = estimate-1.96*std.err, ymax = estimate+1.96*std.err),
              data=rate$TOC, 
              fill = "#1F77B4", alpha = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 1, by = 0.1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.01, suffix = " %")) + # NEW
  labs(title = "Targeting Operator Characteristic", x = "Treated Fraction", y = "") +
  theme_bw() + 
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12))

ggsave("RATE by Gender - Conversion.pdf", width = 10, height = 8)
```

#### By contract valuation (wealth)

```{r}
X.test_tmp <- X.test

X.test_policy_low <- X.test_tmp %>%
  filter(contract_valuation <= median(X.train$contract_valuation))

X.test_policy_high <- X.test_tmp %>%
  filter(contract_valuation > median(X.train$contract_valuation))

policy_low <- average_treatment_effect_test(cf.train, X.test_policy_low)
policy_high <- average_treatment_effect_test(cf.train, X.test_policy_high)

policy_df <- data.frame(
  policy = c("Below Median", "Above Median"),
  estimate = c(policy_low[1], policy_high[1]),
  se = c(policy_low[2], policy_high[2]),
  n = c(nrow(X.test_policy_low), nrow(X.test_policy_high))
)

policy_df$policy <- factor(policy_df$policy, levels = c("Below Median", "Above Median"))

gate_p5 <- ggplot(data = policy_df) +
  geom_col(aes(x = policy, y = estimate, fill = policy)) +
  scale_fill_manual(values = c(`Below Median` = "#A56CBD", `Above Median` = "#1F77B4")) +
  geom_errorbar(aes(x = policy, ymin = estimate - 1.96 * se, ymax = estimate + 1.96 * se), width = 0.1) +
  geom_text(aes(x = 1, y = -0.015), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(policy_df[1,"n"]))) +
  geom_text(aes(x = 2, y = -0.015), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(policy_df[2,"n"]))) +
  geom_text(aes(x = 1, y = 0.025), label = paste0(round(policy_df[1,"estimate"]*100,3), " %")) + # NEW
  geom_text(aes(x = 2, y = 0.025), label = paste0(round(policy_df[2,"estimate"]*100,3), " %")) + # NEW
  ylab("Group ATE") +
  xlab("Contract Valuation") +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("Negative" = expression(pi(x[i]) == 0), "Positive" = expression(pi(x[i]) == 1))) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.01, suffix = " %")) + # NEW
  theme_bw() +
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12)) +
  theme(legend.position = "none")

gate_p5

ggsave("GATE by Contract Valuation - Conversion.pdf", width = 10, height = 8)
```

```{r}
# low -> loss
# high -> gain

X.test_tmp$cate <- cate.test$predictions
X.test_tmp$condition <- W.test
X.test_tmp$converted <- Y.test

# Received gain message and have low value policy (i.e., should have received a loss message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(contract_valuation <= median(X.train$contract_valuation)) %>%
  summarise(n = n(), 
            valuation = mean(contract_valuation),
            base = mean(converted),
            policy_cate = base + policy_low[1]) # estimated conversion rate if they had received a loss message instead

# Received gain message and have high value policy (i.e., should have received a gain message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(contract_valuation > median(X.train$contract_valuation)) %>%
  summarise(n = n(), 
            valuation = mean(contract_valuation),
            base = mean(converted),
            policy_cate = base # nothing to add because they already got the right message
            )
```

```{r}
rate <- rank_average_treatment_effect(cf.test, -1*X.test_tmp$contract_valuation)
rate$estimate
rate$estimate-1.96*rate$std.err
rate$estimate+1.96*rate$std.err

ggplot() +
  geom_line(mapping = aes(x = q, y = estimate), 
            data=rate$TOC, 
            color = "#1F77B4", size = 1) +
  geom_ribbon(mapping = aes(x = q, y = estimate, ymin = estimate-1.96*std.err, ymax = estimate+1.96*std.err),
              data=rate$TOC, 
              fill = "#1F77B4", alpha = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 1, by = 0.1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.01, suffix = " %")) + # NEW
  labs(title = "Targeting Operator Characteristic", x = "Treated Fraction", y = "") +
  theme_bw() + 
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12))

ggsave("RATE by Contract Valuation - Conversion.pdf", width = 10, height = 8)
```

#### By management skills (financial literacy)

```{r}
X.test_tmp <- X.test

X.test_mng_low <- X.test_tmp %>%
  filter(management_skills <= median(X.train$management_skills))

X.test_mng_high <- X.test_tmp %>%
  filter(management_skills > median(X.train$management_skills))

mng_low <- average_treatment_effect_test(cf.train, X.test_mng_low)
mng_high <- average_treatment_effect_test(cf.train, X.test_mng_high)

mng_df <- data.frame(
  mng_skills = c("Below Median", "Above Median"),
  estimate = c(mng_low[1], mng_high[1]),
  se = c(mng_low[2], mng_high[2]),
  n = c(nrow(X.test_mng_low), nrow(X.test_mng_high))
)

mng_df$mng_skills <- factor(mng_df$mng_skills, levels = c("Below Median", "Above Median"))

gate_p5 <- ggplot(data = mng_df) +
  geom_col(aes(x = mng_skills, y = estimate, fill = mng_skills)) +
  scale_fill_manual(values = c(`Below Median` = "#A56CBD", `Above Median` = "#1F77B4")) +
  geom_errorbar(aes(x = mng_skills, ymin = estimate - 1.96 * se, ymax = estimate + 1.96 * se), width = 0.1) +
  geom_text(aes(x = 1, y = -0.015), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(mng_df[1,"n"]))) +
  geom_text(aes(x = 2, y = -0.015), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(mng_df[2,"n"]))) +
  geom_text(aes(x = 1, y = 0.025), label = paste0(round(mng_df[1,"estimate"]*100,3), " %")) + # NEW
  geom_text(aes(x = 2, y = 0.025), label = paste0(round(mng_df[2,"estimate"]*100,3), " %")) + # NEW
  ylab("Group ATE") +
  xlab("Management Skills") +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("Negative" = expression(pi(x[i]) == 0), "Positive" = expression(pi(x[i]) == 1))) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.01, suffix = " %")) + #NEW
  theme_bw() +
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12)) +
  theme(legend.position = "none")

gate_p5

ggsave("GATE by Mng Skills - Conversion.pdf", width = 10, height = 8)
```

```{r}
# low -> loss
# high -> gain

X.test_tmp$cate <- cate.test$predictions
X.test_tmp$condition <- W.test
X.test_tmp$converted <- Y.test

# Received gain message and have high management_skills (i.e., should have received a gain message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(management_skills > median(X.train$management_skills)) %>%
  summarise(n = n(), 
            skills = mean(management_skills),
            base = mean(converted),
            policy_cate = base) # nothing to add because they already got the right message

# Received gain message and low management_skills (i.e., should have received a loss message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(management_skills <= median(X.train$management_skills)) %>%
  summarise(n = n(), 
            skills = mean(management_skills),
            base = mean(converted),
            policy_cate = base + mng_low[1] # estimated conversion rate if they had received a loss message instead
            )
```

```{r}
rate <- rank_average_treatment_effect(cf.test, -1*X.test_tmp$management_skills)
rate$estimate
rate$estimate-1.96*rate$std.err
rate$estimate+1.96*rate$std.err

ggplot() +
  geom_line(mapping = aes(x = q, y = estimate), 
            data=rate$TOC, 
            color = "#1F77B4", size = 1) +
  geom_ribbon(mapping = aes(x = q, y = estimate, ymin = estimate-1.96*std.err, ymax = estimate+1.96*std.err),
              data=rate$TOC, 
              fill = "#1F77B4", alpha = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 1, by = 0.1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.01, suffix = " %")) + #NEW
  labs(title = "Targeting Operator Characteristic", x = "Treated Fraction", y = "") +
  theme_bw() + 
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12))

ggsave("RATE by Mng Skills - Conversion.pdf", width = 10, height = 8)
```

## DV: Saving Amount

```{r}
rm(list=setdiff(ls(), c("relevant_features", "split_ratio", "average_treatment_effect_test", "data_started", "data_finished")))
set.seed(42)
options(scipen=10000)
```

### Fit Causal Forest

Split data into training and test set

```{r}
train <- data_finished %>% filter(year==2021)
test <- data_finished %>% filter(year==2022)
```

Split data into X, Y, W

```{r}
Y.train <- train$paid_amount
W.train <- ifelse(train$condition_3lvl=='loss', 1, 0)
X.train <- train[relevant_features]

Y.test <- test$paid_amount
W.test <- ifelse(test$condition_3lvl=='loss', 1, 0)
X.test <- test[relevant_features]
```

Load or fit model

```{r}
if (file.exists("cf_train_saving_amount.rds")) {
  print("Load CF Train")
  cf.train <- readRDS(file = "cf_train_saving_amount.rds")
} else {
  print("Fit CF Train")
  cf.train <- causal_forest(X = X.train, Y = Y.train, W = W.train,
                            honesty = FALSE,
                            seed = 42)
  saveRDS(cf.train, file = "cf_train_saving_amount.rds")
}
```

###  ATE

```{r}
ate <- average_treatment_effect_test(cf.train, X.test)

print(ate[1])
print(ate[1]-1.96*ate[2])
print(ate[1]+1.96*ate[2])
```

###  CATE

```{r}
cate.test <- predict(cf.train, X.test, estimate.variance = TRUE)
cate.test <- cate.test %>%
  mutate(
    cate_direction = ifelse(cate.test$predictions>=0, 'positive', 'negative') 
  )

summary(cate.test$predictions)
```

```{r}
cate_p1 <- ggplot(cate.test, aes(x = predictions, fill = cate_direction)) +
  geom_histogram(
    binwidth = 100, 
    #alpha = 0.8, 
    color = "white",
    boundary  = 0,              # force a bin edge at exactly zero 
    ) +
  scale_fill_manual(
    name = "Direction",
    values = c(positive = "#1F77B4", negative = "#A56CBD")
  ) +
  labs(
    # x = expression(tau),
    x = "CATE (in €)",
    y = "Frequency"
  ) +
  scale_x_continuous(labels = scales::label_comma()) +
  geom_vline(xintercept=ate[1]) +
  geom_vline(xintercept=c(ate[1]-1.96*ate[2],ate[1]+1.96*ate[2]), linetype="dashed") +
  theme_bw() + 
  theme(legend.position = "none") +
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12)) +
  annotate("text", x = 2000, y = 60, label = sprintf("ATE: %.2f [95%% CI: %.2f, %.2f]", ate[1], ate[1]-1.96*ate[2], ate[1]+1.96*ate[2])) +
  
  annotate("segment", x = 1000, xend = 3000, y = 40, yend = 40, 
           arrow = arrow(length = unit(0.3, "cm")), linewidth = 1) +
  
  annotate("segment", x = -1000, xend = -3000, y = 40, yend = 40, 
           arrow = arrow(length = unit(0.3, "cm")), linewidth = 1) +
  
  annotate("text", x = 2000, y = 30, 
           label = "Respond positively\nto loss-framing", size = 4) +

  annotate("text", x = -2000, y = 30, 
           label = "Respond positively\nto gain-framing", size = 4) +
  
  coord_cartesian(xlim = c(-3100, 3100))

cate_p1
```

### RATE

```{r}
if (file.exists("cf_test_saving_amount.rds")) {
  print("Load CF Test")
  cf.test <- readRDS(file = "cf_test_saving_amount.rds")
} else {
  print("Fit CF Test")
  cf.test <- causal_forest(X = X.test, Y = Y.test, W = W.test,
                          honesty = FALSE,
                          seed = 42)
  saveRDS(cf.test, file = "cf_test_saving_amount.rds")
}
```

```{r}
rate <- rank_average_treatment_effect(cf.test, cate.test$predictions)
rate$estimate
rate$estimate-1.96*rate$std.err
rate$estimate+1.96*rate$std.err
```

```{r}
ggplot(aes(x = q, y = estimate), data=rate$TOC) +
  geom_line(color = "#1F77B4", size = 1) +
  geom_ribbon(aes(ymin = estimate-1.96*std.err, ymax = estimate+1.96*std.err), fill = "#1F77B4", alpha = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 1, by = 0.1)) +
  labs(title = "Targeting Operator Characteristic", x = "Treated Fraction", y = "") +
  theme_bw() + 
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12))

ggsave("RATE - Saving Amount.png", width = 10, height = 8)
```

### GATE

```{r}
X.test_tmp <- X.test
X.test_tmp$cate <- cate.test$predictions

X.test_cate_pos <- X.test_tmp %>% 
  filter(cate > 0) %>% 
  select(-cate)

X.test_cate_neg <- X.test_tmp %>% 
  filter(cate <= 0) %>% 
  select(-cate)

cate_pos <- average_treatment_effect_test(cf.train, X.test_cate_pos)
cate_neg <- average_treatment_effect_test(cf.train, X.test_cate_neg)

cate_sign_df <- data.frame(
  sign = c("Positive", "Negative"),
  estimate = c(cate_pos[1], cate_neg[1]),
  se = c(cate_pos[2], cate_neg[2]),
  n = c(nrow(X.test_cate_pos), nrow(X.test_cate_neg))
)

cate_sign_df <- cate_sign_df %>% arrange(desc(sign))

gate_p1 <- ggplot(data = cate_sign_df) +
  geom_col(aes(x = sign, y = estimate, fill = sign)) +
  scale_fill_manual(values = c(Positive = "#1F77B4", Negative = "#A56CBD")) +
  geom_errorbar(aes(x = sign, ymin = estimate - 1.96 * se, ymax = estimate + 1.96 * se), width = 0.1) +
  scale_y_continuous(labels = scales::label_comma()) +
  geom_text(aes(x = 1, y = -100), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(cate_sign_df[2,"n"]))) +
  geom_text(aes(x = 2, y = 100), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(cate_sign_df[1,"n"]))) +
  geom_text(aes(x = 1, y = -1100), label = round(cate_sign_df[2,"estimate"],2)) +
  geom_text(aes(x = 2, y = 1200), label = round(cate_sign_df[1,"estimate"],2)) +
  ylab("Group ATE") +
  xlab("Causal ML Policy") +
  #ylim(-1800, 1800) +
  # scale_x_discrete(labels = c("Negative" = expression(pi(x[i]) == 0), "Positive" = expression(pi(x[i]) == 1))) +
  scale_x_discrete(labels = c("Negative" = "Not targeted", "Positive" = "Targeted")) +
  theme(legend.position = "none") + 
  theme_bw() +
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12)) +
  theme(legend.position = "none") 

gate_p1
```

Additional CATE-based descriptive statistics

```{r}
detailed_proportions <- X.test_tmp %>% 
  summarise(
    # pos/neg proportions
    prop_positive = mean(cate > 0, na.rm = TRUE),
    prop_negative = mean(cate < 0, na.rm = TRUE),
    
    # interesting thresholds
    prop_gt_500   = mean(cate > 500, na.rm = TRUE),
    prop_gt_1000  = mean(cate > 1000, na.rm = TRUE),
    prop_lt_neg500 = mean(cate < -500, na.rm = TRUE)
  )

detailed_proportions
```

GATE by CATE quartiles

```{r}
X.test_tmp <- X.test
X.test_tmp$cate <- cate.test$predictions
X.test_tmp$cate_quartile <- as.factor(ntile(cate.test$predictions, 4))

X.test_Q1 <- X.test_tmp %>% 
  filter(cate_quartile == 1) %>% 
  select(-cate,-cate_quartile)

X.test_Q2 <- X.test_tmp %>% 
  filter(cate_quartile == 2) %>% 
  select(-cate,-cate_quartile)

X.test_Q3 <- X.test_tmp %>% 
  filter(cate_quartile == 3) %>% 
  select(-cate,-cate_quartile)

X.test_Q4 <- X.test_tmp %>% 
  filter(cate_quartile == 4) %>% 
  select(-cate,-cate_quartile)

gate_q1 <- average_treatment_effect_test(cf.train, X.test_Q1)
gate_q2 <- average_treatment_effect_test(cf.train, X.test_Q2)
gate_q3 <- average_treatment_effect_test(cf.train, X.test_Q3)
gate_q4 <- average_treatment_effect_test(cf.train, X.test_Q4)

gate_quartiles_df <- data.frame(
  quartile = c("Q1", "Q2", "Q3", "Q4"),
  estimate = c(gate_q1[1], gate_q2[1], gate_q3[1], gate_q4[1]),
  se = c(gate_q1[2], gate_q2[2], gate_q3[2], gate_q4[2]),
  sign = c(as.factor(sign(gate_q1[1])), as.factor(sign(gate_q2[1])), as.factor(sign(gate_q3[1])), as.factor(sign(gate_q4[1]))),
  n = c(nrow(X.test_Q1), nrow(X.test_Q2), nrow(X.test_Q3), nrow(X.test_Q4))
)

gate_p2 <- ggplot(data = gate_quartiles_df) +
  geom_col(aes(x = quartile, y = estimate, fill = sign)) +
  scale_fill_manual(values = c("-1" = "#A56CBD", "1" = "#1F77B4")) +
  scale_y_continuous(labels = scales::label_comma()) +
  geom_errorbar(aes(x = quartile, ymin = estimate - 1.96 * se, ymax = estimate + 1.96 * se), width = 0.1) +
 #ylim(-2900, 2900) + 
  ylab("Group ATE") +
  xlab("CATE Quartile") +
  theme_bw() +
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12)) +
  theme(legend.position = "none") 

gate_p2
```

```{r}
plot_grid(cate_p1, gate_p2, gate_p1, labels = c('a', 'b', 'c'), 
          ncol = 1, align = 'h', 
          scale = c(.9, .9, .9)
          )

ggsave("CATE_GATE - Saving Amount.pdf", width = 10, height = 8)
```

### XAI

Variable importance for causal forest (paid amount)

```{r}
vi <- variable_importance(cf.train)
vi_df <- data.frame(vi)
vi_df$variable <- colnames(cf.train$X.orig)

vi_df <- vi_df %>% 
  select("variable", "vi") %>% 
  arrange(desc(vi))

xai_p1 <- ggplot(vi_df, aes(x = reorder(variable, vi), y = vi)) +
  geom_bar(stat = "identity", fill = "#1F77B4") +
  coord_flip() +
  labs(x = "Covariate", y = "Variable Importance") +
  scale_x_discrete(
    labels = function(x) toTitleCase(gsub("_", " ", x))
  ) +
  theme_bw() + 
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12))

xai_p1
```

Partial dependence plots for causal forest mode (paid amount)

```{r}
p_list <- list()
for (var in vi_df$variable) {

  if (var == "gender" | var == "has_company_pension_plan") {
    range <- seq(0, 1, by = 1)
  } else {
    range <- seq(min(X.test[[var]]), max(X.test[[var]]), length.out = 4)    
  }

  range <- seq(min(X.test[[var]]), max(X.test[[var]]), length.out = 4)
  
  results <- lapply(range, function(val) {
    X_mod <- X.test
    X_mod[[var]] <- val
    
    preds <- predict(cf.train, X_mod, estimate.variance = TRUE)
    
    pred_point <- mean(preds$predictions)
    pred_var_mean <- mean(preds$variance.estimates) / nrow(X_mod)
    pred_se <- sqrt(pred_var_mean)
    
    pred_lb <- pred_point - 1.96 * pred_se
    pred_ub <- pred_point + 1.96 * pred_se
    
    data.frame(
      working_with_computers = val,
      pred_point = pred_point,
      pred_lb = pred_lb,
      pred_ub = pred_ub
    )
  })
  
  pdp_df <- do.call(rbind, results)
  names(pdp_df)[1] <- "value"

  p <- ggplot(pdp_df, aes(x = value, y = pred_point)) +
    geom_line(color = "#1F77B4", size = 1) +
    geom_ribbon(aes(ymin = pred_lb, ymax = pred_ub), fill = "#1F77B4", alpha = 0.3) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
    labs(x = str_wrap(toTitleCase(gsub("_", " ", var)), width = 12), y = "CATE") +
    #ylim(y = c(-500, 1200)) +
    theme_bw() + 
    theme(text = element_text(size = 10), axis.text.x=element_text(size=8), axis.text.y=element_text(size=8))

  p_list[[var]] <- p
  
}

p_grid <- plot_grid(plotlist = p_list, ncol = 5, align = 'hv')
p_grid
```

```{r}
plot_grid(
  xai_p1, 
  p_grid,
  labels = c('a', 'b'),
  ncol = 1,
  align = "h",
  scale = c(.9, .9),
  rel_heights = c(0.35, 0.65)
)

ggsave("XAI PDP - Saving Amount Combined.pdf", width = 10, height = 12)
```

SHAP

```{r}
# Prediction wrapper for grf causal forest
pred_fun <- function(object, newdata, ...) {
  predict(object, newdata, ...)$predictions
}

# Compute SHAP values with fastshap
fs <- fastshap::explain(
  object       = cf.train,
  X            = X.test,
  pred_wrapper = pred_fun,
  nsim = 10
)

# Turn fastshap result into a shapviz object
shap_values <- shapviz(fs, X = X.test)

# Global importance plots
#sv_importance(shap_values)
#ggsave("SHAP - PDP Saving Amount.png", width = 10, height = 8)

sv_importance(shap_values, kind = "bee")
ggsave("SHAP - Beeswarm Saving Amount.png", width = 10, height = 8)

```

### Economic Impact

#### Random assignment

Paid amount under randomized assignment

```{r}
X.test_tmp <- X.test
X.test_tmp$cate <- cate.test$predictions
X.test_tmp$condition <- W.test
X.test_tmp$payment <- Y.test

X.test_tmp %>% 
  summarise(n = n(), mean_payment = mean(payment))
```

#### All loss

Paid amount when all participants are assigned to loss-framing

```{r}
X.test_tmp %>% 
  filter(condition == 1) %>%
  summarise(n = n(), mean_payment = mean(payment))
```

#### All gain

Paid amount when all participants are assigned to gain-framing

```{r}
X.test_tmp %>% 
  filter(condition == 0) %>%
  summarise(n = n(), mean_payment = mean(payment))
```

#### By CATE sign (+/-)

Paid amount for optimal assignment to gain- vs. loss-framing

```{r}
# Received gain message and have positive CATE (i.e., should have received a loss message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(cate > 0) %>%
  summarise(n = n(), 
            base = mean(payment),
            delta = base + cate_pos[1]) # estimated saving if they had received a loss message instead

# Received gain message and have negative CATE (i.e., should have received a gain message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(cate <= 0) %>%
  summarise(n = n(), 
            base = mean(payment),
            delta = base # nothing to add because they already got the right message
            )
```

#### By gender

```{r}
X.test_tmp <- X.test

X.test_gender_male <- X.test_tmp %>%
  filter(gender == 0)

X.test_gender_female <- X.test_tmp %>%
  filter(gender == 1)

gender_male <- average_treatment_effect_test(cf.train, X.test_gender_male)
gender_female <- average_treatment_effect_test(cf.train, X.test_gender_female)

gender_df <- data.frame(
  gender = c("Male", "Female"),
  estimate = c(gender_male[1], gender_female[1]),
  se = c(gender_male[2], gender_female[2]),
  n = c(nrow(X.test_gender_male), nrow(X.test_gender_female))
)

gender_df <- gender_df %>% arrange(desc(gender))

gate_p3 <- ggplot(data = gender_df) +
  geom_col(aes(x = gender, y = estimate, fill = gender)) +
  scale_fill_manual(values = c(Male = "#1F77B4", Female = "#A56CBD")) +
  geom_errorbar(aes(x = gender, ymin = estimate - 1.96 * se, ymax = estimate + 1.96 * se), width = 0.1) +
  geom_text(aes(x = 1, y = -500), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(gender_df[2,"n"]))) +
  geom_text(aes(x = 2, y = -500), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(gender_df[1,"n"]))) +
  geom_text(aes(x = 1, y = 800), label = round(gender_df[2,"estimate"],3)) +
  geom_text(aes(x = 2, y = 800), label = round(gender_df[1,"estimate"],3)) +
  ylab("Group ATE") +
  xlab("Gender") +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("Negative" = expression(pi(x[i]) == 0), "Positive" = expression(pi(x[i]) == 1))) +
  theme_bw() +
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12)) +
  theme(legend.position = "none")

gate_p3

ggsave("GATE by Gender - Saving Amount.png", width = 10, height = 8)
```

```{r}
# female -> loss
# male -> gain

X.test_tmp$cate <- cate.test$predictions
X.test_tmp$condition <- W.test
X.test_tmp$converted <- Y.test

# Received gain message and are male (i.e., should have received a gain message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(gender == 0) %>%
  summarise(n = n(), 
            gender = mean(gender),
            base = mean(converted),
            policy_cate = base) # nothing to add because they already got the right message

# Received gain message and are female (i.e., should have received a loss message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(gender == 1) %>%
  summarise(n = n(), 
            gender = mean(gender),
            base = mean(converted),
            policy_cate = base + gender_female[1] # estimated saving if they had received a loss message instead
            )
```

```{r}
rate <- rank_average_treatment_effect(cf.test, X.test_tmp$gender)
rate$estimate
rate$estimate-1.96*rate$std.err
rate$estimate+1.96*rate$std.err

ggplot() +
  geom_line(mapping = aes(x = q, y = estimate), 
            data=rate$TOC, 
            color = "#1F77B4", size = 1) +
  geom_ribbon(mapping = aes(x = q, y = estimate, ymin = estimate-1.96*std.err, ymax = estimate+1.96*std.err),
              data=rate$TOC, 
              fill = "#1F77B4", alpha = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 1, by = 0.1)) +
  labs(title = "Targeting Operator Characteristic", x = "Treated Fraction", y = "") +
  theme_bw() + 
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12))

ggsave("RATE by Gender - Saving Amount.png", width = 10, height = 8)
```

#### By Contract valuation (wealth)

```{r}
X.test_tmp <- X.test

X.test_policy_low <- X.test_tmp %>%
  filter(contract_valuation <= median(X.train$contract_valuation))

X.test_policy_high <- X.test_tmp %>%
  filter(contract_valuation > median(X.train$contract_valuation))

policy_low <- average_treatment_effect_test(cf.train, X.test_policy_low)
policy_high <- average_treatment_effect_test(cf.train, X.test_policy_high)

policy_df <- data.frame(
  policy = c("Below Median", "Above Median"),
  estimate = c(policy_low[1], policy_high[1]),
  se = c(policy_low[2], policy_high[2]),
  n = c(nrow(X.test_policy_low), nrow(X.test_policy_high))
)

policy_df$policy <- factor(policy_df$policy, levels = c("Below Median", "Above Median"))

gate_p5 <- ggplot(data = policy_df) +
  geom_col(aes(x = policy, y = estimate, fill = policy)) +
  scale_fill_manual(values = c(`Below Median` = "#A56CBD", `Above Median` = "#1F77B4")) +
  geom_errorbar(aes(x = policy, ymin = estimate - 1.96 * se, ymax = estimate + 1.96 * se), width = 0.1) +
  geom_text(aes(x = 1, y = -300), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(policy_df[1,"n"]))) +
  geom_text(aes(x = 2, y = -300), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(policy_df[2,"n"]))) +
  geom_text(aes(x = 1, y = 900), label = round(policy_df[1,"estimate"],3)) +
  geom_text(aes(x = 2, y = 900), label = round(policy_df[2,"estimate"],3)) +
  ylab("Group ATE") +
  xlab("Contract Valuation") +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("Negative" = expression(pi(x[i]) == 0), "Positive" = expression(pi(x[i]) == 1))) +
  theme_bw() +
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12)) +
  theme(legend.position = "none")

gate_p5

ggsave("GATE by Contract Valuation - Saving Amount.png", width = 10, height = 8)
```

```{r}
# low -> loss
# high -> gain

X.test_tmp$cate <- cate.test$predictions
X.test_tmp$condition <- W.test
X.test_tmp$converted <- Y.test

# Received gain message and have low value policy (i.e., should have received a loss message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(contract_valuation <= median(X.train$contract_valuation)) %>%
  summarise(n = n(), 
            valuation = mean(contract_valuation),
            base = mean(converted),
            policy_cate = base + policy_low[1]) # estimated saving if they had received a loss message instead

# Received gain message and have high value policy (i.e., should have received a gain message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(contract_valuation > median(X.train$contract_valuation)) %>%
  summarise(n = n(), 
            valuation = mean(contract_valuation),
            base = mean(converted),
            policy_cate = base # nothing to add because they already got the right message
            )
```

```{r}
rate <- rank_average_treatment_effect(cf.test, -1*X.test_tmp$contract_valuation)
rate$estimate
rate$estimate-1.96*rate$std.err
rate$estimate+1.96*rate$std.err

ggplot() +
  geom_line(mapping = aes(x = q, y = estimate), 
            data=rate$TOC, 
            color = "#1F77B4", size = 1) +
  geom_ribbon(mapping = aes(x = q, y = estimate, ymin = estimate-1.96*std.err, ymax = estimate+1.96*std.err),
              data=rate$TOC, 
              fill = "#1F77B4", alpha = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 1, by = 0.1)) +
  labs(title = "Targeting Operator Characteristic", x = "Treated Fraction", y = "") +
  theme_bw() + 
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12))

ggsave("RATE by Contract Valuation - Saving Amount.png", width = 10, height = 8)
```

#### By management skills (financial literacy)

```{r}
X.test_tmp <- X.test

X.test_mng_low <- X.test_tmp %>%
  filter(management_skills <= median(X.train$management_skills))

X.test_mng_high <- X.test_tmp %>%
  filter(management_skills > median(X.train$management_skills))

mng_low <- average_treatment_effect_test(cf.train, X.test_mng_low)
mng_high <- average_treatment_effect_test(cf.train, X.test_mng_high)

mng_df <- data.frame(
  mng_skills = c("Below Median", "Above Median"),
  estimate = c(mng_low[1], mng_high[1]),
  se = c(mng_low[2], mng_high[2]),
  n = c(nrow(X.test_mng_low), nrow(X.test_mng_high))
)

mng_df$mng_skills <- factor(mng_df$mng_skills, levels = c("Below Median", "Above Median"))

gate_p5 <- ggplot(data = mng_df) +
  geom_col(aes(x = mng_skills, y = estimate, fill = mng_skills)) +
  scale_fill_manual(values = c(`Below Median` = "#A56CBD", `Above Median` = "#1F77B4")) +
  geom_errorbar(aes(x = mng_skills, ymin = estimate - 1.96 * se, ymax = estimate + 1.96 * se), width = 0.1) +
  geom_text(aes(x = 1, y = -200), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(mng_df[1,"n"]))) +
  geom_text(aes(x = 2, y = -200), label = paste("N =", scales::label_number(accuracy = NULL, big.mark = ",")(mng_df[2,"n"]))) +
  geom_text(aes(x = 1, y = 800), label = round(mng_df[1,"estimate"],3)) +
  geom_text(aes(x = 2, y = 800), label = round(mng_df[2,"estimate"],3)) +
  ylab("Group ATE") +
  xlab("Management Skills") +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("Negative" = expression(pi(x[i]) == 0), "Positive" = expression(pi(x[i]) == 1))) +
  theme_bw() +
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12)) +
  theme(legend.position = "none")

gate_p5

ggsave("GATE by Mng Skills - Saving Amount.png", width = 10, height = 8)
```

```{r}
# low -> loss
# high -> gain

X.test_tmp$cate <- cate.test$predictions
X.test_tmp$condition <- W.test
X.test_tmp$converted <- Y.test

# Received gain message and have high management_skills (i.e., should have received a gain message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(management_skills > median(X.train$management_skills)) %>%
  summarise(n = n(), 
            skills = mean(management_skills),
            base = mean(converted),
            policy_cate = base) # nothing to add because they already got the right message

# Received gain message and low management_skills (i.e., should have received a loss message)
X.test_tmp %>% 
  filter(condition == 0) %>%
  filter(management_skills <= median(X.train$management_skills)) %>%
  summarise(n = n(), 
            skills = mean(management_skills),
            base = mean(converted),
            policy_cate = base + mng_low[1] # estimated saving if they had received a loss message instead
            )
```

```{r}
rate <- rank_average_treatment_effect(cf.test, -1*X.test_tmp$management_skills)
rate$estimate
rate$estimate-1.96*rate$std.err
rate$estimate+1.96*rate$std.err

ggplot() +
  geom_line(mapping = aes(x = q, y = estimate), 
            data=rate$TOC, 
            color = "#1F77B4", size = 1) +
  geom_ribbon(mapping = aes(x = q, y = estimate, ymin = estimate-1.96*std.err, ymax = estimate+1.96*std.err),
              data=rate$TOC, 
              fill = "#1F77B4", alpha = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 1, by = 0.1)) +
  labs(title = "Targeting Operator Characteristic", x = "Treated Fraction", y = "") +
  theme_bw() + 
  theme(text = element_text(size = 14), axis.text.x=element_text(size=12), axis.text.y=element_text(size=12))

ggsave("RATE by Mng Skills - Saving Amount.png", width = 10, height = 8)
```

# Robustness Checks

## Probit

Extensive margin for 2021 and 2022

```{r}
probit_converted_2021 <- glm(converted ~ condition_3lvl,
                        family=binomial(link='probit'), 
                        data=data_started %>% filter(year==2021))

probit_converted_2022 <- glm(converted ~ condition_3lvl,
                        family=binomial(link='probit'), 
                        data=data_started %>% filter(year==2022))

stargazer(probit_converted_2021, probit_converted_2022,
          type = "text", intercept.bottom = FALSE, single.row = TRUE,
           covariate.labels=c('Constant', 'Received Message: Loss'))
```

## OLS

Extensive margin for 2021 and 2022

```{r}
ols_amount_2021 <- lm(log(paid_amount) ~ condition_3lvl,
                 data=data_finished %>% filter(year==2021))

ols_amount_2022 <- lm(log(paid_amount) ~ condition_3lvl,
                 data=data_finished %>% filter(year==2022))

stargazer(ols_amount_2021, ols_amount_2022,
          type = "text", intercept.bottom = FALSE, single.row = TRUE,
          covariate.labels=c('Constant', 'Received Message: Loss'))
```

## Heckman Two-step selection

Year 2021 (extensive and intensive)

```{r}
selection <- selection(
  converted ~ condition_3lvl +
    historical_payments_cnt,
  
  paid_amount_log ~ condition_3lvl,
  
  data = data_started %>% filter(year == 2021),
  method="2step")

selectionb <- selection
selectionb$param$index$betaO <- selection$param$index$betaS
selectionb$param$index$betaS <- selection$param$index$betaO

stargazer(selectionb, selection, type = "text", intercept.bottom = FALSE, single.row = TRUE, dep.var.labels = 'Two step selection',
                    covariate.labels=c('Constant', 'Received Message: Loss', "Historical Payments"))
```

Year 2022 (extensive and intensive)

```{r}
selection <- selection(
  converted ~ condition_3lvl +
    historical_payments_cnt,
  
  paid_amount_log ~ condition_3lvl,
  
  data = data_started %>% filter(year == 2022),
  method="2step")


selectionb <- selection
selectionb$param$index$betaO <- selection$param$index$betaS
selectionb$param$index$betaS <- selection$param$index$betaO

stargazer(selectionb, selection, type = "text", intercept.bottom = FALSE, single.row = TRUE, dep.var.labels = 'Two step selection',
                              covariate.labels=c('Constant', 'Received Message: Loss', "Historical Payments"))
```